
The basic syntax for an if statement goes like this:

```{r}
#if (logical statement) {
  #do something
#} else {
  #do something else
#}
```
The logical statements should boil down to a boolean; for example:
```{r}
a <- 6
if(a>=10){
  print(a)
}else{
  print("it is not more than 10")
}
```
We can also add else if statements in the middle if we need more options.

```{r}
a <- 15
if (a < 5){
  print('Low')
}else if (a >=5 & a <=10)  {
 print('medium')
}else if (a >= 10 | a ==7){
  print('high')
}else{
  print('hot')
}
  
```

Neat! We can also nest conditionals just like we do in Python.

```{r}
# Print numbers from 1 to 5
for (i in 1:5) {
  print(i)
}
    
```

Neat! Now, let's look at another core part of R programming - Loops!

### Loops

```{r}
for (value in 1:5){
  #paste is a concatenation function
  print(paste(value, "is greater than", value-1))
}
```

We can also do nifty stuff with our vectors, such as apply changes to specific indices of our data.

```{r}
1Vector <- c(1,1,1,1,1,1,1)
#change the first , third, and sixth elements to the value 6
for(i in c(1,3,6)){
  1Vector[i] <- 6
}
Print(1Vector)
1vector
```

We can loop over more than just lists:

```{r}
mat <- matrix(1, nrow= 2, ncol = 2)
for(row in 1:2){
  for(col in 1:2){
    mat[row,col] <- 0
    print(mat)
  }
}
```


```{r}
#Like Python, we have while loops in R:
b <- 6
while(b < 10){
  #increase b by 2
  b <- b + 2
  # print b
  print(b)
}
```

```

## Functions

When should we use functions?

```{r}
our_function <- function(optional_params){
  # do some stuff
  
  
}


```

An example:

```{r}
#start with vector of temperature 
test_temp <- c(16.55678, 26.9420005,32.5910888, 8.580116, 28.003873, 28.5423, 20.976854)                                                                                                 
#Then we want a function that converts from c to f:

change_in_temp <- function(data) {
  temp_fahr <- dat * (9/5) + 32
 
  return(temp_fahr)
}


```

print(change_in_temp(26))
```

This is all cool.. but it didn't do anything. We need to call the function!

```{r}
data_fahr <- change_in_temp(test_temp)
data_fahr
```

Notice that in the background here, we are looping over each item in the array, but we never write a for loop! Keep this in mind, since when you pass in a large data set, the changes made by functions aren't quite immediate - the computer does have to go over each item in your data set, and change them one at a time.

A shortcut of sorts is to use the `apply()` function, which operates very similarly to Python's `map()`. Note that `apply` syntax goes as follows: `apply(X, MARGIN, FUN)` where X is the data set to use, MARGIN is specification for rows or columns, and FUN is the function to apply.

```{r}
test_mat <-  matrix(c(1,1,1,2,2,2), nrow <- 2)
test_mat
apply(test_mat,1 , sum) # this adds the value in the vertical or row order, 1 refers to row and 2 is for colomun
apply(test_mat,2 , sum)
```

It is common to use `mean`, `median`, `sum`, `min`, or `max` with `apply`, but we can also use our user-defined functions as well. Note that apply() only works with matrices.

```{r}
#start with vector of temperature 
test_temp <- c(16.55678, 26.9420005,32.5910888, 8.580116, 28.003873, 28.5423, 20.976854)                                                                                                 
#Then we want a function that converts from c to f:

change_in_temp <- function(dat) {
  temp_fahr <- dat * (9/5) + 32
 
  return(temp_fahr)
}

print(change_in_temp(26))
```
```{r}
test_temp_mat <- matrix(c(16.55678, 26.9420005,32.5910888, 8.580116, 28.003873, 28.5423, 20.976854, 34.7890), nrow <- 2)

apply(test_temp_mat, 1, c_to_f)
```

### Homework

1.  Do 5 CodeWars problems in R. Share your solutions here:
{r}
#one
solution <- function(number) {
  sum <- 0
  
  for (i in 1:(number - 1)) {
    if (i %% 3 == 0 || i %% 5 == 0) {
      sum <- sum + i
    }
  }
  
  return(sum)
}
#TWO

digital_root <- function(n) {

  if (n < 10) {
    return(n)
  }
  
  sum_digits <- sum(as.numeric(strsplit(as.character(n), "")[[1]]))
  

  return(digital_root(sum_digits))
}
#Three
array_diff <- function(a, b) {
  
  b_set <- unique(b)
  
  result <- a[!(a %in% b_set)]
  
  return(result)
}
#Four 
duplicate_count <- function(text) {
  text <- tolower(text)
  

  char_count <- list()
  
  for (char in strsplit(text, "")[[1]]) {
  
    if (!grepl("[a-z0-9]", char)) {
      next
    }
    

    if (is.null(char_count[[char]])) {
      char_count[[char]] <- 1
    } else {
      char_count[[char]] <- char_count[[char]] + 1
    }
  }
  

  duplicates_count <- sum(unlist(char_count) > 1)
  
  return(duplicates_count)
}


    ```{r}

```

3.  **Mario Challenge:** In the classic Super Mario games, Mario ends each level by ascending a flight of stairs and leaping off the top to arrive at the castle at the end. (See [here](https://youtu.be/-avspZlbOWU?t=55) for details). We can imagine that each of the bricks of the stairs is a character X. Create a function that prints out a set of stairs. For example:

    If the function is called marioStairs() and we pass in the value 4, we would expect to see stairs with a length of 4 and a height of 4:

    X

    XX

    XXX

    XXXX




  
```{r}
#You may feel more comfortable figuring this out in Python first, then translating into R later.
  for (i in 0:3) {
    for (j in 0:i) {
      last("X ")
    }
    last("\n \n")
  }

```
Create a function that checks an input of type int and returns fizz if the number is a multiple of 3, buzz if the number is a #multiple of 5, and fizzbuzz if the number is a multiple of 3 and 5.

```{r}
fizz_Orbuzz <- function(n) {
    result <- c()
    for (i in 1:n) {
      if (i %% 3 == 0 ) {
        result <- c(result, "Fizz")
      } else if (i %% 3 == 0 && i %% 5 == 0) {
        result <- c(result, "Fizzbuzz")
      } else if (i %% 5 == 0) {
        result <- c(result, "Buzz")
      } else {
        result <- c(result, as.character(i))
      }
    }
    return(result)
}
fizz_Orbuzz(12)

```